---
title:  "Sustainable Tourism: Data Analysis of the Spanish Tourism Industry (1995-2022)"

author: "Name: Kwadwo ('Ko-jo') Gyampo. Email: fkgyampo@gmail.com"

date:   "August, 2024"

output: pdf_document

toc: FALSE 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# *1.0 Introduction* 
According to United Nations World Tourism Organisation (UNWTO), in 1950 there were 
25 million international tourism arrivals, which increased to 1.3 billion in 2017.
The international tourism sector is expected to grow by 3.3% annually, until 2030,
representing an expected 1.8 billion tourists crossing international borders.

It is projected, as the Human Development Index (HDI) improves across the world
resulting in increased incomes per capita or per household, seasonal peaks in international 
travel may become a mainstay of the global tourism economy. However, the growth of tourism
has led to concerns of mass travel leading to overtourism at selected travel destinations in 
Europe causing a schism between local residents and tourists. 

In Spain, several protests by local residents and activist groups were observed at 
favorite tourism destinations particularly in the cities of Barcelona, Malaga and Seville 
as well as the Balearic Islands made up of the Canary Islands, Formentera, Majorca, 
Menorca and Ibiza. 

The general concerns of protesting groups and local residents were hinged on the cascading 
effects of mass tourism and overtourism resulting in an increased cost of living and 
a reduced quality of life for residents. 

The UNWTO asserts that in 2023, "there were over 85 million tourists in Spain compared 
to a population of 47 million", further adding impetus to the concerns of overtourism and 
bringing into focus the carrying capacity of popular tourism destinations in the European
country.

The goal of the data analysis of Spanish Tourism spanning the years 1995 to 2022, 
is to investigate the key concerns of overtourism by analysing key inbound travel metrics. 

## *1.1 Research Objectives* 
The objectives of the data analysis project are as follows:
  a. Assess the trend of inbound travel visitors to Spain over 25 years. 
  
  b. Investigate the regions of inbound travel to Spain spanning 25 years.
  
  c. Determine the travel expenditure and travel purpose for inbound travelers. 
  
  d. Assess the occupancy rates and availability of accommodation units for inbound visitors. 
  
  e. Establish the summary statistical metrics for inbound tourists. 
  
  f. Outline the employment rates for direct and downstream hospitality industries engaged in 
     Spanish tourism.


## *1.2 Research Question*

The main research question with respect to the data analysis is as follows:
  a. Is Overtourism a fact or a myth in the Spanish tourism industry ?
  

## *1.3 Methodology* 
The methodology for the analysis was carried out as follows:
  a. The database for UNWTO which serves as a renowned global travel repository 
     for UN registered nations was searched for credible and verifiable data on Spain. 
     
  b. The global dataset publicly available on the unwto website (www.unwto.org) was downloaded. 
  
  c. Further, all specific sub-datasets relevant to Spain were extricated from the 
     global repository. 
  
  d. To ensure data integrity, the sub-datasets were renamed and aligned to the file names
     of the global dataset. 
  
  e. The datasets were pre-processed in MS excel to remove extra spaces (rows and columns) 
     and irrelevant symbols without altering the cell contents. 
  
  f. The cleaned and processed files were renamed as secondary files, resaved as 
     comma separated value (CSV) files and exported to RStudio. 
  
  g. Imported CSV files into RStudio, were further wrangled and transformed with relevant 
     codes and syntaxes prior to analyses. 
     
  h. Final data cleaning, transformation and all visualizations were carried out in 
     RStudio in line with the research objectives. 
  
  i. Finally, a report was knitted in Rmarkdown to display all codes, visualizations and 
     conclusion of the study. 

## *1.4 Discussion - Exploratory Data Analysis* 
A preliminary investigative analysis of the dataframes, generated after cleaning, 
transformation and wrangling of the data files, was carried out to glean preliminary 
statistical trends. 

The Exploratory Data Analysis (EDA) focused on statistical summaries, percentage variations, 
cumulative summations, means as well as peaks (maximum) and troughs (minimum) values of trends. Percentage distributions for all datasets were conducted to observe trends and glean important observations. 

## *1.5 Conclusion*
The following conclusions were drawn:
a. Data analysis of the Spanish tourism industry displays a water shed moment for the 
    Spanish tourism industry in 1999. The analysis indicates the pivotal year for tourism 
    (1999) experienced a significant growth for accommodation units especially for the number 
    of bed-places. The number of bed-places sequentially experienced a robust year-on-year 
    growth from 1999 and peaked in 2022 at 2 million units.

b. Despite the tourism down-turn of 2020 possibly attributed to the Covid-19 travel 
   shutdowns, the sharp rise in occupancy rates in 2021 to 2022 attests to the allure and                 attractiveness of Spanish holiday destinations to inbound tourist arrivals. 


c. Data analysis of inbound tourism arrivals for the period under review, displayed a 
   chronological year-on-year growth for groups of visitor arrivals.

d. Analysis of regional tourism arrivals to Spain showed an overwhelming patronage of 
   Spanish tourism destinations by arrivals from Europe. It is fair to assume that travel 
   policies including a unified European Schengen visa programme facilitates a fluid 
   and seamless intra-European travel. 

e. Tourists contributed more than 80million dollars to the Spanish economy in 2018.

f. Most tourists surveyed during the period under review indicated two main travel 
  purposes: business and personal. A majority of tourists indicated the purpose of 
  travel as personal.

g. It was observed that the food and beverage industry generated about 51% of total 
   employment in tourism. This was followed by the passenger transportation and other 
   tourism industries. In third place for tourism employment by industry sub-sector 
   was accommodation services and hotels. 

4.1 Overtourism – A fact or a myth 
The available data and metrics established were not sufficient to investigate the case 
of overtourism. To establish the existence of overtourism in the Spanish tourism industry, 
metrics such as described below must be established and validated by data analysis:

a. Mass tourism: This term refers to the large numbers of tourists visiting a destination, 
  often resulting in overcrowding and strain on local resources.

b. Tourist saturation: This describes the point at which a destination becomes overwhelmed 
   by the number of tourists, leading to negative impacts on the local environment, economy, 
   and community.
   
c. Carrying capacity: This refers to the maximum number of tourists that a destination can 
   accommodate without experiencing negative impacts on the environment, infrastructure, 
   and local community.
   
d. Tourism overload: This phrase describes the situation in which a destination is unable 
   to cope with the number of tourists, leading to congestion, overcrowding, and strain 
   on local resources.
   
e. Unsustainable tourism: This term refers to tourism practices that are not environmentally, 
   socially, or economically sustainable, often resulting in negative impacts on the local 
   community and environment.


```

```
***

***
```
```
```
2.0 Data Analysis
Install packages for analysis of datasets
  * tidyverse  
  * patchwork  
  * viridis  
  * RColorBrewer  
  * ggmap
```

```{r, echo=FALSE, Include=FALSE,eval=TRUE}
# 2.0 install packages 
#install.packages("tidyverse")
library(tidyverse)

#install.packages("patchwork")
library(patchwork)

#install.packages("RColorBrewer")
library(RColorBrewer)

#install.packages("ggmap")
library(ggmap)
```

```
2.1 Load all data files

  a. tourism industry (accommodation) csv file  
  
  b. inbound tourism arrivals csv file  
  
  c. inbound tourism by region csv file  
  
  d. inbound tourism expenditure csv file  
  
  e. inbound tourism purpose of travel csv file  
  
  f. tourism employment csv file
  
  g. protests across spain csv file

```

```
3.0 Tourism Industry Accommodation Analysis

3.1 Load all data files

```

```{r,echo=FALSE,include=FALSE,results='hide',warning=FALSE,message=FALSE}
#3.1 Load all data files 
  #a. 1. tourism industry csv file
  tourism.accommodation <- read.csv(file.choose(),header=TRUE)

  #b. 2. inbound tourism arrivals csv file
  inbound.spain <- read.csv(file.choose(),header=TRUE)
  
  #c. 3. inbound tourism by region csv file
  inbound.tourism.regions <- read.csv(file.choose(),header=TRUE)
  
  #d. 4. inbound tourism expenditure csv file
  inbound.expenditure <- read.csv(file.choose(),header=TRUE)
  
  #e. 5. inbound tourism purpose csv file
  inbound.tourism.purpose <- read.csv(file.choose(),header=TRUE)
  
  #f. 6. Tourism Employment csv file
  tourism.employment <- read.csv(file.choose(),header=TRUE)
  
  #g. 7. Protests Across Spain csv file
  protests <- read.csv(file.choose(),header=TRUE)
```


```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#3.2 select relevant variables from loaded dataframe 
spain.tourism.accommodation <- tourism.accommodation %>% 
                                      select("Visitors.accommodation",c("X1995":"X2022"))
                               
spain.tourism.accommodation

```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#3.3 rename variables 
spain.tourism.accomm.1 <- spain.tourism.accommodation %>% 
                                  rename("1995" = "X1995",
                                         "1996" = "X1996",
                                         "1997" = "X1997",
                                         "1998" = "X1998",
                                         "1999" = "X1999",
                                         "2000" = "X2000",
                                         "2001" = "X2001",
                                         "2002" = "X2002",
                                         "2003" = "X2003",
                                         "2004" = "X2004",
                                         "2005" = "X2005",
                                         "2006" = "X2006",
                                         "2007" = "X2007",
                                         "2008" = "X2008",
                                         "2009" = "X2009",
                                         "2010" = "X2010",
                                         "2011" = "X2011",
                                         "2012" = "X2012",
                                         "2013" = "X2013",
                                         "2014" = "X2014",
                                         "2015" = "X2015",
                                         "2016" = "X2016",
                                         "2017" = "X2017",
                                         "2018" = "X2018",
                                         "2019" = "X2019",
                                         "2020" = "X2020",
                                         "2021" = "X2021",
                                         "2022" = "X2022")
spain.tourism.accomm.1
```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#3.4 Coerce years to numeric variables
spain.tourism.accomm.2 <- spain.tourism.accomm.1 %>% 
                   mutate(1995 == as.numeric("1995","1996","1997","1998","1999","2000","2001",
                                             "2002","2003","2004","2005","2006","2007","2008",
                                             "2009","2010","2011","2012","2013","2014","2015",
                                             "2016","2017","2018","2019","2020","2021","2022"))
           
spain.tourism.accomm.2

```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#3.5 filter valid values from dataframe

spain.tourism.accomm.2 <- spain.tourism.accomm.2 %>% 
                    filter(!is.na(Visitors.accommodation),na.rm=TRUE,
                           !is.na(1995),na.rm=TRUE,
                           !is.na(1996),na.rm=TRUE,
                           !is.na(1997),na.rm=TRUE,
                           !is.na(1998),na.rm=TRUE,
                           !is.na(1999),na.rm=TRUE,
                           !is.na(2000),na.rm=TRUE,
                           !is.na(2001),na.rm=TRUE,
                           !is.na(2002),na.rm=TRUE,
                           !is.na(2003),na.rm=TRUE,
                           !is.na(2004),na.rm=TRUE,
                           !is.na(2005),na.rm=TRUE,
                           !is.na(2006),na.rm=TRUE,
                           !is.na(2007),na.rm=TRUE,
                           !is.na(2008),na.rm=TRUE,
                           !is.na(2009),na.rm=TRUE,
                           !is.na(2010),na.rm=TRUE,
                           !is.na(2011),na.rm=TRUE,
                           !is.na(2012),na.rm=TRUE,
                           !is.na(2013),na.rm=TRUE,
                           !is.na(2014),na.rm=TRUE,
                           !is.na(2015),na.rm=TRUE,
                           !is.na(2016),na.rm=TRUE,
                           !is.na(2017),na.rm=TRUE,
                           !is.na(2018),na.rm=TRUE,
                           !is.na(2019),na.rm=TRUE,
                           !is.na(2020),na.rm=TRUE,
                           !is.na(2021),na.rm=TRUE,
                           !is.na(2022),na.rm=TRUE)
spain.tourism.accomm.2

```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#3.6 transform dataframe structure with pivot_longer function
spain.tourism.accomm.3 <- spain.tourism.accomm.2 %>% 
                        pivot_longer(cols=`1995`:`2022`,
                           names_to = "year",
                           values_to = "total.visitors"
                          )
spain.tourism.accomm.3
```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#3.7 select relevant variables from transformed dataframe 
spain.tourism.accomm.4 <- spain.tourism.accomm.3 %>% 
                      select(Visitors.accommodation,year,total.visitors)
spain.tourism.accomm.4
```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#3.8 retrieve levels of Visitors.accommodation 

  fct_count(spain.tourism.accomm.4$Visitors.accommodation)
  
```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#3.9 plot a line graph for categorical variable visitors.accommodation 
  
#3.9.1 group each level for visitors accommodation 
      spain.tourism.accomm.4$group_var <- spain.tourism.accomm.4$Visitors.accommodation
      
     (spain.tourism.accomm.4.group_var <- spain.tourism.accomm.4$Visitors.accommodation)
      
```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#3.9.2 However, occupancy rates are percentages. remove these levels for homogenisation of variables
  
      d <- spain.tourism.accomm.4 %>% 
              filter(!Visitors.accommodation %in% c("Occupancy rate / bed-places","Occupancy rate / rooms")) 
      d
```


```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hold'}
#3.9.3 plot final graph - add arguments for final reporting
  
    d <- d %>% 
        ggplot(aes(x = year, y = Visitors.accommodation, group = group_var, color = group_var, 
                   linetype = group_var, size = group_var)) + 
        geom_point(aes(group=group_var,size=0.5,shape=group_var))+
        geom_line(linetype="solid",size=1.2)+
        theme_bw() + 
        theme(axis.text.x = element_text(angle = 90, hjust = 1), 
              axis.title.x = element_text(size = 8, face = "bold"), 
              axis.title.y = element_text(size = 8, face = "bold"), 
              legend.title = element_text(size = 8, face = "bold"), 
              legend.position = "bottom") + 
        labs(title="Spanish Tourism:Number of Accommodation Units have Increased since 1998",
             x="Year",
             y="Accommodation Units",
             color = "Accommodation for Visitors",
             caption= "Source:www.unwto.org",
             legend.position="bottom")+
        scale_color_brewer(palette = "Set1") + 
        scale_linetype_manual(values = c("solid", "solid", "solid","solid")) + 
        #scale_size_manual(values = c(1.5, 1.5, 1.5,1.5))+
        guides(linetype = "none", size = "none",shape="none")
d
```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#3.10 plotting occupancy rates in percentages separately
      d.1 <- spain.tourism.accomm.4 %>% 
                  filter(Visitors.accommodation %in% c("Occupancy rate / bed-places","Occupancy rate / rooms")) 
      d.1

```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hold'}
d.1 <- d.1 %>% 
        ggplot(aes(x = year, y = Visitors.accommodation, group = group_var, color = group_var, 
                   linetype = group_var, size = group_var)) + 
        geom_point(aes(group=group_var,size=0.5,shape=group_var))+
        geom_line(linetype="solid",size=1.2)+
        theme_bw() + 
        theme(axis.text.x = element_text(angle = 90, hjust = 1), 
              axis.title.x = element_text(size = 8, face = "bold"), 
              axis.title.y = element_text(size = 8, face = "bold"), 
              legend.title = element_text(size = 8, face = "bold"), 
              legend.position = "bottom") + 
        labs(title="Spanish Tourism:Accommodation Occupancy Rates have Surged to Pre-covid Heights",
             x="Year",
             y="Percentage Occupancy (%)",
             color = "Occupancy for Visitors",
             caption= "Source:www.unwto.org") +
        scale_color_brewer(palette = "Set2") + 
        scale_linetype_manual(values = c("solid", "solid")) + 
        #scale_size_manual(values = c(1.5, 1.5))+
        guides(linetype = "none", size = "none",shape="none")
d.1
  
```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hold'}
#3.11 Dashboard of d and d.1
  
  (d | d.1)+
    plot_annotation()+
    theme()
```
***


***

```
# *4.0 Inbound Arrivals Analysis*

```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
# 4.1 select relevant variable columns 
      inbound.spain.1 <- inbound.spain %>% 
                                  select("Year",c("X1995":"X2022"))
      inbound.spain.1 
```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
# 4.2 tidy, clean and transform dataset 
      inbound.spain.2 <- inbound.spain.1 %>% 
                              rename("Arrivals" = "Year",
                                      "1995" = "X1995",
                                      "1996" = "X1996",
                                      "1997" = "X1997",
                                      "1998" = "X1998",
                                      "1999" = "X1999",
                                      "2000" = "X2000",
                                      "2001" = "X2001",
                                      "2002" = "X2002",
                                      "2003" = "X2003",
                                      "2004" = "X2004",
                                      "2005" = "X2005",
                                      "2006" = "X2006",
                                      "2007" = "X2007",
                                      "2008" = "X2008",
                                      "2009" = "X2009",
                                      "2010" = "X2010",
                                      "2011" = "X2011",
                                      "2012" = "X2012",
                                      "2013" = "X2013",
                                      "2014" = "X2014",
                                      "2015" = "X2015",
                                      "2016" = "X2016",
                                      "2017" = "X2017",
                                      "2018" = "X2018",
                                      "2019" = "X2019",
                                      "2020" = "X2020",
                                      "2021" = "X2021",
                                      "2022" = "X2022")
inbound.spain.2
```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
# 4.3 select relevant variables and omit missing values 
inbound.spain.3 <- inbound.spain.2 %>% 
                        filter(!is.na("Arrivals"),na.rm=TRUE,
                               !is.na(1995),na.rm=TRUE,
                               !is.na(1996),na.rm=TRUE,
                               !is.na(1997),na.rm=TRUE,
                               !is.na(1998),na.rm=TRUE,
                               !is.na(1999),na.rm=TRUE,
                               !is.na(2000),na.rm=TRUE,
                               !is.na(2001),na.rm=TRUE,
                               !is.na(2002),na.rm=TRUE,
                               !is.na(2003),na.rm=TRUE,
                               !is.na(2004),na.rm=TRUE,
                               !is.na(2005),na.rm=TRUE,
                               !is.na(2006),na.rm=TRUE,
                               !is.na(2007),na.rm=TRUE,
                               !is.na(2008),na.rm=TRUE,
                               !is.na(2009),na.rm=TRUE,
                               !is.na(2010),na.rm=TRUE,
                               !is.na(2011),na.rm=TRUE,
                               !is.na(2012),na.rm=TRUE,
                               !is.na(2013),na.rm=TRUE,
                               !is.na(2014),na.rm=TRUE,
                               !is.na(2015),na.rm=TRUE,
                               !is.na(2016),na.rm=TRUE,
                               !is.na(2017),na.rm=TRUE,
                               !is.na(2018),na.rm=TRUE,
                               !is.na(2019),na.rm=TRUE,
                               !is.na(2020),na.rm=TRUE,
                               !is.na(2021),na.rm=TRUE,
                               !is.na(2022),na.rm=TRUE)
inbound.spain.3

```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
# 4.4 transform structure of final dataframe with pivot_longer() function 
inbound.spain.4 <- inbound.spain.3 %>% 
                        pivot_longer(cols=`1995`:`2022`,
                                     names_to = "year",
                                     values_to = "total.visitors")
inbound.spain.4
```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
# 4.5 group levels of "Arrivals"
(inbound.spain.4$group_var <- inbound.spain.4$Arrivals)


    #or to group using mutate
    #inbound.spain.4 <- inbound.spain.4 %>% 
                            #mutate(group_var = Arrivals)
    #inbound.spain.4
```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hold'}
# 4.6 final plot - line chart - inbound tourism group arrivals
inbound.spain.4 %>% 
  ggplot(aes(x = year, y = total.visitors, group = group_var, color = group_var, 
             linetype = group_var, size = group_var)) + 
  geom_point(aes(group=group_var,shape=group_var,size=0.5))+
  geom_line(size=1.2) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        axis.title.x = element_text(size = 10, face = "bold"), 
        axis.title.y = element_text(size = 10, face = "bold"), 
        legend.title = element_text(size = 10, face = "bold"), 
        legend.position = "bottom") + 
  labs(title   = "Spain:Inbound tourism arrivals have peaked in 2022",
       x       = "Year", 
       y       = "Total Arrivals (thousands)", 
       color   = "Groups of Arrivals",
       caption = "Source:www.unwto.org.com") + 
  scale_color_brewer(palette = "Set1") + 
  scale_linetype_manual(values = c("solid", "solid", "solid")) + 
  #scale_size_manual(values = c(1.5, 1.5, 1.5))+
  guides(linetype = "none", size = "none",shape="none")
```

***


***
```
# *5.0 Inbound Arrivals by regions Analysis* 
```


```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
# 5.1 select column variables 

inbound.tourism.regions.1 <- inbound.tourism.regions %>% 
                                    select("year",c("X1995":"X2022"))
inbound.tourism.regions.1

```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#5.2 rename column variables 
inbound.tourism.regions.2 <- inbound.tourism.regions.1 %>% 
                                    rename("Regions" = "year",
                                           "1995" = "X1995",
                                           "1996" = "X1996",
                                           "1997" = "X1997",
                                           "1998" = "X1998",
                                           "1999" = "X1999",
                                           "2000" = "X2000",
                                           "2001" = "X2001",
                                           "2002" = "X2002",
                                           "2003" = "X2003",
                                           "2004" = "X2004",
                                           "2005" = "X2005",
                                           "2006" = "X2006",
                                           "2007" = "X2007",
                                           "2008" = "X2008",
                                           "2009" = "X2009",
                                           "2010" = "X2010",
                                           "2011" = "X2011",
                                           "2012" = "X2012",
                                           "2013" = "X2013",
                                           "2014" = "X2014",
                                           "2015" = "X2015",
                                           "2016" = "X2016",
                                           "2017" = "X2017",
                                           "2018" = "X2018",
                                           "2019" = "X2019",
                                           "2020" = "X2020",
                                           "2021" = "X2021",
                                           "2022" = "X2022")
inbound.tourism.regions.2
```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#5.3 select relevant variables 
inbound.tourism.regions.3 <- inbound.tourism.regions.2 %>% 
                                     filter(!is.na(Regions),na.rm=TRUE,
                                            !is.na(1995),na.rm=TRUE,
                                            !is.na(1996),na.rm=TRUE,
                                            !is.na(1997),na.rm=TRUE,
                                            !is.na(1998),na.rm=TRUE,
                                            !is.na(1999),na.rm=TRUE,
                                            !is.na(2000),na.rm=TRUE,
                                            !is.na(2001),na.rm=TRUE,
                                            !is.na(2002),na.rm=TRUE,
                                            !is.na(2003),na.rm=TRUE,
                                            !is.na(2004),na.rm=TRUE,
                                            !is.na(2005),na.rm=TRUE,
                                            !is.na(2006),na.rm=TRUE,
                                            !is.na(2007),na.rm=TRUE,
                                            !is.na(2008),na.rm=TRUE,
                                            !is.na(2009),na.rm=TRUE,
                                            !is.na(2010),na.rm=TRUE,
                                            !is.na(2011),na.rm=TRUE,
                                            !is.na(2012),na.rm=TRUE,
                                            !is.na(2013),na.rm=TRUE,
                                            !is.na(2014),na.rm=TRUE,
                                            !is.na(2015),na.rm=TRUE,
                                            !is.na(2016),na.rm=TRUE,
                                            !is.na(2017),na.rm=TRUE,
                                            !is.na(2018),na.rm=TRUE,
                                            !is.na(2019),na.rm=TRUE,
                                            !is.na(2020),na.rm=TRUE,
                                            !is.na(2021),na.rm=TRUE,
                                            !is.na(2022),na.rm=TRUE)
inbound.tourism.regions.3
```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#5.4 restructure dataframe using pivot_longer function 
inbound.tourism.regions.4 <- inbound.tourism.regions.3 %>% 
                                      pivot_longer(cols=`1995`:`2022`,
                                                   names_to = "Years",
                                                   values_to = "Regional Arrivals")
inbound.tourism.regions.4

```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#5.5 group character levels of Regions using group_var function 

#5.5.1 using group_var

inbound.tourism.regions.4$group_var <- inbound.tourism.regions.4$Regions

```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hold'}
#5.6 plot - regional distribution of inbound tourism arrivals
inbound.tourism.regions.plota <- inbound.tourism.regions.4 %>% 
                                      filter(!`Regional Arrivals` %in% c("Total")) %>% 
                                      filter(!group_var %in% c("Total")) %>% 
                                      ggplot(aes(x=Years,y=`Regional Arrivals`,group=group_var,color=group_var,
                                                 linetype = group_var, size = group_var)) +
                                      geom_point(aes(group=group_var,shape=group_var,size=0.5))+
                                      geom_line(size=1.4) +
                                      theme_bw() + 
                                      theme(axis.text.x = element_text(angle = 90, hjust = 1), 
                                            axis.title.x = element_text(size = 10, face = "bold"), 
                                            axis.title.y = element_text(size = 10, face = "bold"), 
                                            legend.title = element_text(size = 10, face = "bold"), 
                                            legend.position = "bottom") + 
                                      labs(title   = "Regional Arrivals:Europe tops other regions in patronage of Spanish tourism",
                                            x      = "Year", 
                                            y      = "Total Visitors (thousands)", 
                                           color   = "Regional Arrivals",
                                           caption = "Source:www.unwto.org") + 
                                      scale_color_brewer(palette = "Set1") + 
                                      scale_linetype_manual(values = c("solid", "solid", "solid","solid","solid","solid","solid")) + 
                                      #scale_size_manual(values = c(1.5, 1.5, 1.5,1.5,1.5,1.5,1.5))+
                                      guides(linetype = "none", size = "none",shape="none")

inbound.tourism.regions.plota

```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hold'}
#5.7 Regional distribution of tourism arrivals. Disaggregation by region
inbound.tourism.regions.plotb <- inbound.tourism.regions.4 %>% 
                                    filter(!`Regional Arrivals` %in% c("Total")) %>% 
                                    filter(!group_var %in% c("Total")) %>% 
                                    ggplot(aes(x=Years,y=`Regional Arrivals`,group=group_var,color=group_var,
                                                linetype = group_var, size = group_var)) + 
                                    geom_line(size=1.2) +
                                    facet_wrap(~group_var)+
                                    theme(panel.grid.major.x = element_blank(),
                                          panel.grid.major.y = element_line(size = 0.5, linetype = "solid",color="grey"),
                                          panel.background = element_blank(),
                                          legend.key = element_rect(colour = "black")) + 
                                    theme(axis.text.x = element_text(angle = 90, hjust = 1), 
                                          axis.title.x = element_text(size = 10, face = "bold"), 
                                          axis.title.y = element_text(size = 10, face = "bold"), 
                                          legend.title = element_text(size = 10, face = "bold"), 
                                          legend.position = "right") + 
                                    labs(title = "Regional distribution of tourism arrivals in Spain",
                                      subtitle = "(Disaggregation by region)",
                                            x  = "Year", 
                                            y  = "Total Visitors(thousands)", 
                                         color = "Regional Arrivals",
                                         caption = "Source:www.unwto.org") + 
                                    scale_color_brewer(palette = "Set1") + 
                                    scale_linetype_manual(values = c("solid", "solid", "solid","solid","solid","solid","solid")) + 
                                    scale_size_manual(values = c(1.5, 1.5, 1.5,1.5,1.5,1.5,1.5))+
                                    guides(linetype = "none", size = "none",shape = "none",color="none")
                                                                     
inbound.tourism.regions.plotb

```

```
***

***

```

```
6.0 Inbound Tourism Expenditure Analysis

```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#6.1 select relevant variables 
inbound.expenditure.1 <- inbound.expenditure %>% 
                                  select("Year",c("X1995":"X2022"))
inbound.expenditure.1

```


```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#6.2 rename variables 
inbound.expenditure.2 <- inbound.expenditure.1 %>% 
                                  rename("Tourism Expenditure"="Year",
                                         "1995" = "X1995",
                                         "1996" = "X1996",
                                         "1997" = "X1997",
                                         "1998" = "X1998",
                                         "1999" = "X1999",
                                         "2000" = "X2000",
                                         "2001" = "X2001",
                                         "2002" = "X2002",
                                         "2003" = "X2003",
                                         "2004" = "X2004",
                                         "2005" = "X2005",
                                         "2006" = "X2006",
                                         "2007" = "X2007",
                                         "2008" = "X2008",
                                         "2009" = "X2009",
                                         "2010" = "X2010",
                                         "2011" = "X2011",
                                         "2012" = "X2012",
                                         "2013" = "X2013",
                                         "2014" = "X2014",
                                         "2015" = "X2015",
                                         "2016" = "X2016",
                                         "2017" = "X2017",
                                         "2018" = "X2018",
                                         "2019" = "X2019",
                                         "2020" = "X2020",
                                         "2021" = "X2021",
                                         "2022" = "X2022")
inbound.expenditure.2

```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#6.3 transform df structure
inbound.expenditure.3 <- inbound.expenditure.2 %>% 
                                pivot_longer(cols=`1995`:`2022`,
                                             names_to = "Years",
                                             values_to = "Total Expenditure")
inbound.expenditure.3

```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#6.4 group levels 
inbound.expenditure.4 <- inbound.expenditure.3 %>% 
                                mutate(group_var = `Tourism Expenditure`)
inbound.expenditure.4
```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hold'}
#6.5 plot - Inbound expenditure in Spain
inbound.expenditure.5 <- inbound.expenditure.4 %>% 
                                filter(!`Tourism Expenditure` %in% c("Tourism expenditure in the country","Passenger transport")) %>% 
                                ggplot(aes(x=Years,y=`Total Expenditure`,group=group_var,color=group_var,
                                           size=group_var,line = group_var)) +
                                geom_line(aes(color=group_var))+
                                theme(axis.text.x=element_text(angle=90,hjust=1))+
                                theme_bw()
inbound.expenditure.5  
```


```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hold'}
#6.6 final plot - Inbound travel expenditure by tourists
inbound.expenditure.6 <- inbound.expenditure.4 %>% 
                                filter(!`Tourism Expenditure` %in% c("Tourism expenditure in the country","Passenger transport")) %>% 
                                ggplot(aes(x=Years, y=`Total Expenditure`, fill=group_var)) + 
                                geom_bar(position="dodge", stat="identity",color="black",alpha=0.5) +
                                coord_flip()+
                                theme_bw() + 
                                theme(axis.text.x  = element_text(angle = 90, hjust = 1), 
                                      axis.title.x = element_text(size = 8, face = "bold"), 
                                      axis.title.y = element_text(size = 8, face = "bold"), 
                                      legend.title = element_text(size = 8, face = "bold"), 
                                      legend.position = "right") + 
                                labs(title = "Spain:Estimated travel expenditure by inbound tourists",
                                     x     = "Year", 
                                     y     = "Total Expenditure($millions)", 
                                     fill  = "Travel Expenditure") + 
                                scale_fill_brewer(palette = "Set1")
inbound.expenditure.6

```

****


****

```
7.0 Inbound Tourism Purpose of Travel Analysis  

```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#7.1 rename variables 
inbound.tourism.purpose.1 <- inbound.tourism.purpose %>% 
                                    rename("Purpose"="X",
                                           "1995" = "X1995",
                                           "1996" = "X1996",
                                           "1997" = "X1997",
                                           "1998" = "X1998",
                                           "1999" = "X1999",
                                           "2000" = "X2000",
                                           "2001" = "X2001",
                                           "2002" = "X2002",
                                           "2003" = "X2003",
                                           "2004" = "X2004",
                                           "2005" = "X2005",
                                           "2006" = "X2006",
                                           "2007" = "X2007",
                                           "2008" = "X2008",
                                           "2009" = "X2009",
                                           "2010" = "X2010",
                                           "2011" = "X2011",
                                           "2012" = "X2012",
                                           "2013" = "X2013",
                                           "2014" = "X2014",
                                           "2015" = "X2015",
                                           "2016" = "X2016",
                                           "2017" = "X2017",
                                           "2018" = "X2018",
                                           "2019" = "X2019",
                                           "2020" = "X2020",
                                           "2021" = "X2021",
                                           "2022" = "X2022")
inbound.tourism.purpose.1

```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#7.2 select relevant variables 
inbound.tourism.purpose.2 <- inbound.tourism.purpose.1 %>% 
                                        select("Purpose",c("1995":"2022"))
inbound.tourism.purpose.2

```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#7.3 restructure df with pivot_longer function
inbound.tourism.purpose.3 <- inbound.tourism.purpose.2 %>%
                                        pivot_longer(cols=`1995`:`2022`,
                                                     names_to = "Years",
                                                     values_to = "Travel purpose")
inbound.tourism.purpose.3  

```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hide'}
#7.4 group purpose 
inbound.tourism.purpose.4 <- inbound.tourism.purpose.3 %>% 
                                      mutate(group_var = Purpose)
inbound.tourism.purpose.4

```

```{r,echo=TRUE,include=TRUE,eval=TRUE,results='hold'}
#7.5 plot
inbound.tourism.purpose.5 <- inbound.tourism.purpose.4 %>% 
                                      filter(!Purpose %in% c("Total")) %>% 
                                      ggplot(aes(x=Years,y=`Travel purpose`,fill=group_var))+
                                      geom_bar(position="stack",stat="identity",alpha=0.8,color="black")+
                                      coord_flip()+
                                      theme_bw()+
                                      theme(axis.text.x=element_text(angle=90,hjust=1))+
                                      labs(title = "Spain:Travel purpose by inbound tourists",
                                           x     = "Years",
                                           y     = "Tourists (thousands)",
                                           fill  = "Travel purpose",
                                           caption="Source:www.unwto.org")

inbound.tourism.purpose.5  
```


```
8.0 Tourism Employment by Industries

```


```{r,echo=FALSE,include=FALSE,eval=TRUE,results='hide'}
employment.spain <- read.csv(file.choose(),header=TRUE)
```


```{r,echo=FALSE,include=TRUE,results='hold'}
#8.1 Annual Employment Generated by Tourism
employment.dist <- employment.spain %>% 
                              select(Year,Total) %>% 
                              ggplot(aes(x=Year,y=Total))+
                              geom_bar(position="dodge",stat="identity",fill="red",color="black",alpha=0.5)+
                              #coord_cartesian(xlim(c(1996,2022)))+
                              coord_flip()+
                              theme_bw()+
                              labs(title = "Spain:Annual Employment Generated by Tourism",
                                   x     = "Year",
                                   y     = "Total Employment(thousands)",
                                   caption = "Source:www.unwto.org")+
                              scale_fill_brewer(palette = "Set1")
employment.dist                                         

```



```{r,echo=TRUE,include=TRUE,results='hide'}
#8.2 percentage distribution by industry 
# the dataframe is named 'employment.spain'
# Melt the dataframe to convert the percentage columns into a single column
employment.spain_melted <- reshape2::melt(employment.spain, id.vars = "Year", 
                            variable.name = "Category", 
                            value.name = "Percentage")
employment.spain_melted
```

```{r,echo=TRUE,include=TRUE,results='hide'}

df.employment.spain_melted <- employment.spain_melted %>% 
                                filter(!Category %in% c("Total")) %>% 
                                group_by(Year,Category,Percentage) %>% 
                                count(n=n()) %>% 
                                arrange(desc(Percentage))
df.employment.spain_melted


```

```{r,echo=TRUE,include=TRUE,results='hold'}
# 8.3 Plot the bar graph using ggplot2
employment.spain_melted %>% 
                        filter(!Category %in% c("Total")) %>% 
                          ggplot(aes(x = Year, y = Percentage, fill = Category)) + 
                            geom_bar(stat = "identity", position = "stack") + 
                            coord_flip()+
                            
                            labs(x = "Year", 
                                 y = "Percentage", 
                                 title = "Spain:Employment in Percentages by Year"
                                ) + 
                            theme(legend.position = "bottom")+
                            theme_bw()+
                            scale_fill_brewer(palette="Set1")
```




```{r,echo=FALSE,include=TRUE,results='hold'}
#8.4 Tourism Employment by Percentages 
df.employment.spain_melted %>% 
                      #filter(!Category %in% c("Total")) %>% 
                      ggplot(aes(x = fct_reorder(Category,Percentage), y = Percentage)) + 
                      geom_bar(stat = "identity", position = "stack") + 
                      coord_flip()+
                      #scale_y_continuous(limits = c(0,100))+
                      labs(x = "Tourism Industry", y = "Percentage Employment", title = "Spain: Tourism Industry Employment in Percentages by Year"
                           ) + 
                      theme_bw()+
                      scale_fill_brewer(palette="Set1")+
                       guides(x = "none")
                     
```

```{r,echo=FALSE,include=TRUE,results='hold'}

#8.5 Tourism Employment by Percentages (by Category)
df.employment.spain_melted %>% 
                      #filter(!Category %in% c("Total")) %>% 
                      ggplot(aes(x = fct_reorder(Category,Percentage), y = Percentage,fill=Category)) + 
                      geom_bar(stat = "identity", position = "stack") + 
                      coord_flip()+
                      #scale_y_continuous(limits = c(0,100))+
                      labs(x = "Tourism Industry", y = "Percentage Employment", title = "Spain: Tourism Industry Employment in Percentages by Year"
                           ) + 
                      theme_bw()+
                      scale_fill_brewer(palette="Set1")+
                       guides(x = "none",fill = "none")
                     
```

****


****


```{r}
#9.0  plot map of region of Spain
       #install.packages("ggmap")
       #library(ggmap)
       
       #install.packages("mapdata")
       #library(mapdata)
```

```{r,echo=FALSE,include=FALSE,results='hide'}
(Spain <- map_data("world", region = c("Spain")))
```

```{r,results='asis',fig.width=8,fig.height=6}
Spain <-  map_data("world",region=c("Spain"),subregion=c("Formentera","Ibiza","Majorca","Minorca"),complete.cases=TRUE)

       
Spain.map <- ggplot(Spain, aes(x=long, y=lat, group=group, fill=subregion)) +
         geom_polygon(color="black",linewidth=0.8) +
         scale_fill_brewer(palette="Set1")+
         theme_light()+
         labs(title = "Spain Tourism Hotspots:Balearic Island Archipelagos",
              fill = "Balearic Islands")+
         theme(axis.title.x = element_text(size = 8))+
         theme(axis.title.y = element_text(size = 8))
Spain.map
```

```{r}
#protests
        
anti.tourism.protests <- protests %>% 
                          ggplot(aes(x=fct_reorder(City,Number.of.Protesters),y=Number.of.Protesters))+
                          geom_bar(aes(fill=City),position="dodge",stat="identity")+
                          geom_text(aes(label=Number.of.Protesters), size=4, vjust=-0.5,hjust=0.5) +
                          scale_fill_brewer(palette="Set1")+
                          theme_bw()+
                          guides(fill="none")+
                          labs(title = "Anti-tourism protests across cities in Spain",
                               x     = "Cities",
                               y     = "Number of Protesters",
                             caption = "Source:www.wikipedia.org")
anti.tourism.protests 

```

